HyperLogLog: Exponentially Bad in Adversarial Settings

### 2022-02-05 docs/HLL_exp_bad/main.tex
update results text

We first focus on a classic HLL implementation as per the original HLL paper~\cite{hll}, namely the implementation of~\cite{clahll}.
We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}; again the attacks do not depend on the specifics of what fixed hash function is used.
%(note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
We create an HLL sketch using $n=8$ and hence $m=256$ buckets and, for attacks in the setting of an initially non-empty target HLL sketch, initialise it with 1,000 random items representing honest users' data.

@@

As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from ${\mathcal{X}}$, the adversary has to choose what to insert from a subset of 250,000 random items that are given to it.

@@

First, when attacking a sketch already containing 1000 random items, we can see that the adversary is on average able to insert about half of the $2^{20}$ items in our starting set in scenario S2, and about 83\% in scenario S4. The cardinality estimate is, on average, increased slightly in S2 (from 1,001 to 1,025) and remains at the same starting value (1,007 on average) in S4. In both cases, we inserted more than half a million items whilst increasing the cardinality estimate by about $0.2\%$!  These results align well with our theoretical analysis.  
%In scenario S1 we run the attack designed for an initially empty target HLL sketch. It performs well in practice, for example, increasing the HLL cardinality estimate from 100 to 126????
%( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$)
%Here $0.83 \approx (1-\frac{1}{2\cdot H_c})$ with $H_c\approx\frac{1000}{256}=3.9$, so the results are also consistent with our theoretical analysis.
%% That analysis was now updated -- does it still match?
Second, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly a rounded version of the Hit Counting estimator $m \cdot \ln (m/(m-B))$ after inserting a fraction $\frac{B}{m}$ of the items, as predicted by our analysis. For example, in scenario S1 with $B=128$, the final cardinality estimate is 177 (in agreement with the formula) while with $B=1$ in scenarios S2, S4, the final cardinality estimate is just 1 (again, as predicted by the formula).

>>>>

We first focus on a classic HLL implementation as per the original HLL paper~\cite{hll}, namely the implementation of~\cite{clahll}.
We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}; again the attacks do not depend on the specifics of what fixed hash function is used.
%(note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
We create an HLL sketch using $n=8$ and hence $m=256$ buckets and, for attacks in the setting of an initially non-empty target HLL sketch, initialise it with $t \cdot m$ log $m$ random items representing honest users' data.

@@

As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mount our attack in their own setting: instead of picking from ${\mathcal{X}}$, the adversary has to choose what to insert from a subset of 250,000 random items that are given to it.

@@

First, when attacking a sketch already containing $t \cdot m$ log $m$ random items, we can see that the adversary is on average able to insert a fraction $1-1/2^t$ of the $2^{20}$ items in our starting set in scenario S2, and a fraction $1-H_M^{-1}$ in scenario S4. For example for $t = 1$, the adversary can insert half of the items in S2, and about 83\% in S4. The cardinality estimate is, on average, increased slightly in S2 (from 1,411 to 1,430) and remains at the same starting value (1,416 on average) in S4. In both cases, we inserted more than half a million items whilst increasing the cardinality estimate by about $1\%$!  These results align well with our theoretical analysis.  
%In scenario S1 we run the attack designed for an initially empty target HLL sketch. It performs well in practice, for example, increasing the HLL cardinality estimate from 100 to 126????
%( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$)
%Here $0.83 \approx (1-\frac{1}{2\cdot H_c})$ with $H_c\approx\frac{1000}{256}=3.9$, so the results are also consistent with our theoretical analysis.
%% That analysis was now updated -- does it still match?
Second, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly a rounded version of the Hit Counting estimator $m \cdot \ln (m/(m-B))$ after inserting a fraction $\frac{B}{m}$ of the items, as predicted by our analysis. For example, in scenario S1 with $B=128$, the final cardinality estimate is 177 (in agreement with the formula) while with $B=1$ in scenarios S2, S4, the final cardinality estimate is just 1 (again, as predicted by the formula).
